\subsection{Quantisation and Coding Efficiency} \label{sec:quantisation}

Image X is filtered with two filters of the same shape but different lengths - 3, $h_3 = \frac{1}{4}[1{\ }2{\ }1]$ and 5, $h_5 = \frac{1}{16}[1{\ }4{\ }6{\ }4{\ }1]$. The  filtered images are then quantised under 2 compression schemes, the constant step size scheme and the constant MSE scheme. The constant step size scheme maintains a constant quantisation step size between the pyramid layers to give the same RMS error as the direct quantisation of the original image at a step size of 17, while the constant MSE scheme maintains a constant quantisation step size ratio between the layers to give the same RMS error. The respective step size scales for different filters and compression schemes can be found in Figure \ref{fig:step sizes}. The MSE step size ratios for the 3 and 5 tap filters are obtained from the impulse response analysis and found to be $ss_3 = [1.000 {\ } 0.667 {\ } 0.364 {\ } 0.186 {\ } 0.094 {\ } 0.047 {\ } 0.023 {\ } 0.006]$ and $ss_5 = [1.000 {\ } 0.914 {\ } 0.506 {\ } 0.259 {\ } 0.130 {\ } 0.065 {\ } 0.033 {\ } 0.006]$ respectively. The respective compression ratios are shown in Figure \ref{fig:compression ratios} and the relevant reconstructed images are shown in Figure \ref{fig:lap3} and \ref{fig:lap5} respectively. \\

From Figure \ref{fig:compression ratios}, it can be observed that for the constant step size scheme, both filters have the best compression ratio at $1.389$ and $1.335$ respectively, when a 2 layer pyramid is used. By considering the visual effects of the reconstructed images in Figure \ref{fig:const_step_lap3} and \ref{fig:const_step_lap5}, together with the compression ratios, we conclude that the optimal number of pyramid layers for this scheme is 2. For the MSE scheme, it is observed from Figure \ref{fig:compression ratios} that the compression ratio remains relatively similar for a pyramid with 2 or more layers at $1.549$ and $1.480$ respectively for the 3 and 5-tap filters. For the 3-tap filter, the best compression ratio is obtained at $1.550$ with 4 layers. Considering the decoded images in Figure \ref{fig:const_mse_lap3}, we conclude that layers between 2 and 5 gives similar visual performances, thus an optimal layer choice would be 4. For the  5-tap filter, the best compression ratio is at $1.485$ with 5 layers. The decoded images for pyramid of layers 3 to 7 are visually similar, hence an optimal choice of pyramid layers is 5. \\

From Figure \ref{fig:compression ratios}, it can be observed that the performance of the constant step size scheme decreases with more than 2 layers. This is expected as with more layers, more pixels are required to be encoded, resulting in a smaller compression ratio. The constant MSE scheme is found to have a plateauing performance for more than 2 layers, and generally performs better than the constant step size scheme. This is because the step size ratios is such that the quantisation step size is larger for a layer with many image pixels. This means that the bits required to be encoded remains relatively similar with more layers, giving a similar compression ratio, unlike the constant step scheme, which has an increase in number of bits with more layers. It is also found that the 3-tap filter generally has a better performance compared to the 5-tap filter, as a longer filter length results in more definition and higher energy in the high-pass images, hence more bits are required to encode the images, giving in a poorer performance.
